---
title: "HW1 Number1"
author: "Aaron Coates"
date: "4/17/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("~/Documents/GitHub/MMSS_311_2")
sick <- read.csv("/Users/aaroncoates/Downloads/sick_data.csv")

library(dplyr)
library(ggplot2)
library(glmnet)
library(broom)
library(aod)
```

First, I turn positive results to "1" and negative results to "0".
```{r}
sick$result.d <- sapply(sick$result, function(x){
  if(x == "Positive"){
    1
  } else {
    0
  }})
```

This is the OLS regression.
```{r}
regOLS <- glm(result.d ~ temp + bp, data=sick) %>% broom::tidy()
regOLS
for (i in 1:1000) {sick$predictionOLS[i] <- 
  regOLS[1, 2] + regOLS[2, 2]*sick$temp[i] + regOLS[3,2]*sick$bp[i]}
```

Above, I generated predicted values using the OLS regression. I summarize the results below.
```{r}
(sum(with(sick, result.d==1 & predictionOLS>=.5)) + 
   sum(with(sick, result.d==0 & predictionOLS<=.5)))/1000
```

So, the model is accurate 96.4% of the time.
Now, I will rearrange variables to find the line of best fit.
```{r}
intOLS <- (regOLS[1, 2] -.5)/(-regOLS[3, 2]) 
slopeyOLS <- regOLS[2, 2]/(-regOLS[3,2])
intOLS
slopeyOLS
```

I plot the results below.
```{r}
plotOLS <- ggplot(data = sick, aes(x = temp, y = bp, col=result)) + 
  geom_point() + geom_abline(intercept = intOLS[1, 1], 
  slope = slopeyOLS[1, 1]) +
  labs(title ="OLS", x = "Temperature", y = "Blood Pressure")
```

```{r, echo=FALSE}
plotOLS
```

Now, I will perform the logistic regression.
```{r}
sick$result <- factor(sick$result)
logit <- glm(result ~ temp + bp, data = sick, 
             family = "binomial") %>% broom::tidy()
logit
```

Now, I will generate predicted values and then summarize the results.
```{r}
for (i in 1:1000) {sick$predictionLOGIT[i] <- 
  logit[1, 2] + logit[2, 2]*sick$temp[i] + logit[3,2]*sick$bp[i]}
```

```{r}
(sum(with(sick, result.d==1 & predictionLOGIT>=.5)) + 
   sum(with(sick, result.d==0 & predictionLOGIT<=.5))) / 1000
```

So, the model is 99.2% accurate. Now, I will find the line of best fit by rearranging coefficients.
```{r}
intLOG <- (logit[1, 2] -.5)/(-logit[3, 2]) 
slopeyLOG <- logit[2, 2]/(-logit[3,2])
intLOG
slopeyLOG
```

```{r}
plotLOG <- ggplot(data = sick, aes(x = temp, y = bp, col=result)) + 
  geom_point() + geom_abline(intercept = intLOG[1, 1], 
  slope = slopeyLOG[1, 1]) + labs(title ="Logit", 
  x = "Temperature", y = "Blood Pressure")
```

```{r, echo=FALSE}
plotLOG
```

